Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-03-29 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Graph Algoriths in MapReduce}

  RMAT generation, pagerank, CC, triangle-finding, Luby, SSSP, other?
  how many variants of something like CC do we want to include?
    I would vote for just one
  discuss possible inefficiency due to a giant CC in CC algorithm

>>>>

\section{Graph Algoriths in MapReduce}
\label{sec:graph}

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT graphs ~\cite{RMAT} are recursively generated
graphs with power-law degree distributions.  They are commonly used to
represent web and social networks.  The user specifices 6 parameters
which define the graph: the number of vertices $N$ and edges $M$, and
4 parameters $a$, $b$, $c$, $d$ which sum to 1.0 and are discussed
below.  The algorithm in Figure \ref{fig:rmat} generates $N_z = MN$
unique non-zero entries in a sparse $NxN$ matrix $A$, where each entry
$A_{ij}$ represents an edge between graph vertices $(V_i,V_j)$.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for R-MAT graph generation.}

 \label{fig:rmat}
\end{figure}

In the map() operation, each of $P$ processors is tasked with
generating a $1/P$ fraction of the desired edges.  A single random
$i,j$ edge is computed recursively as follows.  Pick a random quadrant
of the $A$ matrix with relative probabilities $a$, $b$, $c$, and $d$.
Treat the chosen quadrant as a sub-matrix and select a random quadrant
within it, in the same manner.  Repeat this process $n$ times where $N
= 2^n$.  At the end of the recursion, the final ``quadrant'' is
non-zero matrix element $A_{ij}$.

The map() will often generate some small number of duplicate edges.
The collate() and reduce() operations remove the duplicates.  The
entire map-collate-reduce sequence is repeated until the number of
resulting key/value pairs $N_{kv}$ equals $N_z$.  For reasonably
sparse graphs this typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other 3, a highly skewed distribution
results.  Variants of the above algorithm can be used when $N$ is not
a power-of-two, to generate graphs with weighted edges (assign a
random numeric value to the $A_{ij}$ edge), graphs without self edges
(require $i != j$), or graphs with undirected edges (require $i < $j).

Discussion of PageRank and its matvec kernel.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for PageRank vertex ranking.}

 \label{fig:rmat}
\end{figure}

Discussion of triangle finding algorithm.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for triangle finding.}

 \label{fig:rmat}
\end{figure}

Discuss connected component identification algorithm.

Do we need differnt variants of CC?

Could just do one, discuss possible inefficiency due to giant CC.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for connected component identification.}

 \label{fig:rmat}
\end{figure}

Discussion of Luby algorithm for maximally independent sets.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for R-MAT graog generation.}

 \label{fig:rmat}
\end{figure}

Discussion of SSSP.

Discuss new optimization in SSSP?

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{MapReduce algorithm for single-source shortest path (SSSP)
 determination.}

 \label{fig:rmat}
\end{figure}

### 2010-03-29 paper/outcore.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

\item {Pass 4) A set file is read and the key/value data for each of
its KV pairs is copied into the appropriate location in the page of
KMV pairs, using information stored in the HT.  When complete, the KMV
page is written to disk.  This pass eventually reads all set files,
entailing a final read of all KV pairs.  It also writes all KMV pairs
to disk.  If each KV pair has a unique key, the volume of KMV output
is roughly the same as that of the KV input.

>>>>

\item (Pass 4) A set file is read and the key/value data for each of
its KV pairs is copied into the appropriate location in the page of
KMV pairs, using information stored in the HT.  When complete, the KMV
page is written to disk.  This pass eventually reads all set files,
entailing a final read of all KV pairs.  It also writes all KMV pairs
to disk.  If each KV pair has a unique key, the volume of KMV output
is roughly the same as that of the KV input.

### 2010-03-29 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

\centerline{Keywords: MapReduce, message-passing, MPI, graph
algorithms, RMAT matrices}

@@

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions which operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of Terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized RMAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

>>>>

\centerline{Keywords: MapReduce, message-passing, MPI, graph
algorithms, R-MAT matrices}

@@

We describe a parallel library written with message-passing (MPI)
calls that allows algorithms to be expressed in the MapReduce
paradigm.  This means the calling program does not need to include
explicit parallel code, but instead provides ``map'' and ``reduce''
functions which operate independently on elements of a data set
distributed across processors.  The library performs needed data
movement between processors.  We describe how typical MapReduce
funtionality can be implemented in an MPI context, and also in an
out-of-core manner for data sets that do not fit within the aggregate
memory of a parallel machine.  Our motivation for creating this
library was to enable graph algorithms to be written as MapReduce
operations, allowing processing of Terabyte-scale data sets.  We
outline MapReduce versions of several such algorithms: vertex ranking
via PageRank, triangle finding, connected component identification,
Luby's algorithm for maximally independent sets, and single-source
shortest-path calculation.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized R-MAT matrices in
parallel; a MapReduce version of this operation is also described.
Performance and scalability results for the various algorithms are
presented for varying size graphs on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

### 2010-03-29 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@479 3fc257bb-8550-0410-81b3-a5c103da83f7
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@484 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Performance Results}

  ideally, for each problem:
    define chosen characteristics of graph, e.g. edges/vertex
    choose 3 sizes:
      small = fits on 1 proc)
      medium = out-of-core on 1 proc, but not on P procs
      huge = impressive size
  show performance plots for each size running on 1 to P procs
    would like to see kink in plots as go out-of-core
    be able to judge and explain basic scaling results
  what machine: Odin or Tbird or RStorm?
    something that maps to a vanilla cluster
  where we have data:
    show multiple machines
    alternate non-MR algorithms
      discuss alternate CC algorithms via Trilinos and hybrid
    alternate MR implementations (Hadoop)

>>>>

\section{Performance Results}
\label{sec:results}

In this section, we present performance results for the MapReduce
graph algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.

The benchmarks were run on a medium-sized Linux cluster.  Describe
Odin.

We ran each of the algorithms on 3 artificial R-MAT graphs of
different sizes, each on a varying number of processors.  The {\it
small} problem size (around 1M edges) was a problem that could be run
on a single processor.  The {\it medium} problem size (around 1B
edges) could be run on a few processors.  The {\it large} problem size
(around 1T edges) required most of the machine to run.

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.

Discuss R-MAT generation times.

Discuss PageRank performance.  Include Trilinos and other
algorithms and machines?

Discuss triangle finding times.

Discuss connected component identification times.

Discuss Luby maximally independent sets times.

Discuss SSSP times.
